Analyzing Security-Cost Trade-off for a \\ Fully Homomorphic Encryption Scheme

### 2013-06-16 Report/article.tex
j

\subsection{Cost of Cryptosytem}
 In the SV scheme, performing any operation on encrypted bits requires  \texttt{KeyGen}, \texttt{Encrypt}, \texttt{Add},\texttt{ Mult}, \texttt{Recrypt}, \texttt{Decrypt} and hence we measure their performance in terms of run time. These measuremens are done by varying the parameter $N$ that determines the degree of all intermediate polynomials in the cryptostystem and keeping all other parameters fixed. The measurements in this section were taken on the machine M2.

@@

\autoref{fig:image_f2} measures execution time for \texttt{KeyGen} for fixed values of parameters: $N=8$ and $\eta = 2^{384}$. It shows a variation from 0.2 s to 14.41 s with high density in the interval [0.2s, 2s].

@@

1. \textbf{XOR of an $n$-bit sequence}\\
2. \textbf{Sum of :}  two $n$ bit sequences  \newline 
3. \textbf{Majority bit of an $n$-bit sequence} \\
4. \textbf{Product of two $n\times n$ boolean matrices} \\
5. \textbf{Sorting of $n$ sequences of $nbit$ using :} \newline\noindent
\phantom{x}\hspace{3ex} a) Insertion sort \phantom{x}\hspace{2ex} b) Bitonic sort \\
	\phantom{x}\hspace{3ex} 	     c) Odd-Even Merge sort

@@

\textbf{XOR of $n$-bits :} \texttt{XOR} of $n$-bits requires $n-1$ \texttt{XOR} gates. \newline \newline
\textbf{Sum of two $n$-bit sequences :} Sum of two $n$-bits sequences is performed using $n$ \texttt{full-adder} circuits (provided as \texttt{fhe\_fulladd(...)} function in the Scarab library). The cost of a full adder circuit is $2$ \texttt{XOR} to calculate the sum and $2$ \texttt{AND}, $1$ \texttt{XOR} and $1$ \texttt{OR} for the carry. Hence the total complexity is $3n$ \texttt{AND}, $5n$ \texttt{XOR}. \newline \newline
\textbf{Majority of $n$-bits :} Majority bit is evaluated by obtaining the sum of the bits and then returning the result of the comparison of the sum and $n/2$. This comparison is done using algorithm \autoref{Code:algo}. The operation is interesting in the sense that if the same program is evaluated on the user's side, he has to store the sum of the bits, hence  needs $\mathrm{\lfloor\log_{2}n \rfloor +1}$ bits of memory. However, if he performs the operation remotely on a cloud, he only requires a constant number of bits to retrieve the result. \newline \newline
\textbf{Product of two $n\times n$ matrices :} This requires $n^3$ \texttt{AND} gates and \texttt{$n^2(n-1)$} \texttt{XOR} gates. \newline \newline
\textbf{Sorting :} Several straight-line programs for sorting have have been proposed in literature \cite{dk}, \cite{Batcher:1968:SNA:1468075.1468121}. These programs use a comparator gate : a logical gate with two inputs and two outputs, that computes the minimum and maximum of two $n$ bit sequences. Algorithm \autoref{Code:algo} implements such a comparator gate. We do not need to perform any transformation for these algorithms. The algorithms used for Bitonic and Odd-Even merge sort are provided in the appendix. They accept $n$ which are powers of $2$.
\newline \newline
\textbf{Insertion sort :} When allowing for parallel comparators, bubble sort and insertion sort are identical. Hence $n(n-1)/2$ comparator gates are used for insertion sort.\newline \newline
\textbf{Bitonic Sort :} The standard algorithm uses $n\mathrm{log}(n)·(\mathrm{log}(n)+1)/4$ comparator gates. \newline \newline
\textbf{Odd-Even Merge sort :} Odd-Even merge sort requires  $n(\mathrm{log}^{2}(n)-1)/2 + 1 $ comparator gates.\newline
We note that the sorting programs used do not increase the size of the circuit operating on encrypted input bit sequences.

\section{Experimental evaluation \\ and results}
\label{Sec:Eval}

\textbf{XOR of $n$ bits}
\begin{figure}[!H]%on ouvre l'environnement figure
\centering
\includegraphics[width=7cm, height=6cm]{f4.pdf} 
\caption{Wall time for XOR} 
\label{fig:image_sf7} %l'étiquette pour faire référence à cette image
\end{figure}

\textbf{Sum of two $n$ bit sequences}
\begin{figure}[!H]%on ouvre l'environnement figure
\centering
\includegraphics[width=6cm]{f8.pdf} 
\caption{Wall time for sum} 
\label{fig:image_sf7} %l'étiquette pour faire référence à cette image
\end{figure}

\textbf{Majority of $n$ bits}

\textbf{Product of two $n\times n$ matrices}

\begin{figure}[!H]%on ouvre l'environnement figure
\centering
\includegraphics[width=9cm, height=6cm]{f9.pdf} 
\caption{Wall time for matrix product} 
\label{fig:image_sf7} %l'étiquette pour faire référence à cette image
\end{figure}

>>>>

\subsection{Cost of Cryptosytem}
 In the SV scheme, performing any operation on encrypted bits requires  \texttt{KeyGen}, \texttt{Encrypt}, \texttt{Add},\texttt{ Mult}, \texttt{Recrypt}, \texttt{Decrypt} and hence we measure their performance in terms of run time. These measuremens are done by varying the parameter $N$ that determines the degree of all intermediate polynomials in the cryptostystem and keeping all other parameters fixed. The measurements in this section were taken on the machine M2. We can infer that the cost of these functions increase with the increase of the the degree  polynomials.

@@

We can see that the size of the public key is increasing lineary with $n$ such that $N=2^n$ 
\autoref{fig:image_f2} measures execution time for \texttt{KeyGen} for fixed values of parameters: $N=8$ and $\eta = 2^{384}$. It shows a variation from 0.2 s to 14.41 s with high density in the interval [0.2s, 2s].

@@

1. \textbf{XOR of an $n$-bit sequence}\\
3. \textbf{Majority bit of an $n$-bit sequence} \\
2. \textbf{Sum of :}  two unsigned integers of $32$ bits  \newline 
2. \textbf{Sum of :}  two unsigned integers  \newline 
4. \textbf{Product of two $n\times n$ boolean matrices} \\
5. \textbf{Sorting of $n$ sequences of $nbit$ using :} \newline\noindent
\phantom{x}\hspace{3ex} a) Insertion sort \phantom{x}\hspace{2ex} b) Bitonic sort \\
	\phantom{x}\hspace{3ex} 	     c) Odd-Even Merge sort

@@

\textbf{XOR of $n$-bits :} \texttt{XOR} of $n$-bits requires $n-1$ \texttt{XOR} gates. \newline \newline

\textbf{Majority of $n$-bits :} Majority bit is evaluated by obtaining the sum of the bits and then returning the result of the comparison of the sum and $n/2$. This comparison is done using algorithm \autoref{Code:algo}. The operation is interesting in the sense that if the same program is evaluated on the user's side, he has to store the sum of the bits, hence  needs $\mathrm{\lfloor\log_{2}n \rfloor +1}$ bits of memory. However, if he performs the operation remotely on a cloud, he only requires a constant number of bits to retrieve the result. \newline \newline

\textbf{Sum of two unsigned integers of $32$ bits :} Sum of two $n$-bits sequences is performed using $n$ \texttt{full-adder} circuits (provided as \texttt{fhe\_fulladd(...)} function in the Scarab library). The cost of a full adder circuit is $2$ \texttt{XOR} to calculate the sum and $2$ \texttt{AND}, $1$ \texttt{XOR} and $1$ \texttt{OR} for the carry. Hence the total complexity is $3n$ \texttt{AND}, $5n$ \texttt{XOR}. \newline \newline

\textbf{Sum of two unsigned integers :} Sum of two $n$-bits sequences is performed using $n$ \texttt{full-adder} circuits (provided as \texttt{fhe\_fulladd(...)} function in the Scarab library). The cost of a full adder circuit is $2$ \texttt{XOR} to calculate the sum and $2$ \texttt{AND}, $1$ \texttt{XOR} and $1$ \texttt{OR} for the carry. Hence the total complexity is $3n$ \texttt{AND}, $5n$ \texttt{XOR}. \newline \newline

\textbf{Product of two $n\times n$ matrices :} This requires $n^3$ \texttt{AND} gates and \texttt{$n^2(n-1)$} \texttt{XOR} gates. \newline \newline
\textbf{Sorting :} Several straight-line programs for sorting have have been proposed in literature \cite{dk}, \cite{Batcher:1968:SNA:1468075.1468121}. These programs use a comparator gate : a logical gate with two inputs and two outputs, that computes the minimum and maximum of two $n$ bit sequences. Algorithm \autoref{Code:algo} implements such a comparator gate. We do not need to perform any transformation for these algorithms. The algorithms used for Bitonic and Odd-Even merge sort are provided in the appendix. They accept $n$ which are powers of $2$.
\newline \newline
\textbf{Insertion sort :} When allowing for parallel comparators, bubble sort and insertion sort are identical. Hence $n(n-1)/2$ comparator gates are used for insertion sort.\newline \newline
\textbf{Bitonic Sort :} The standard algorithm uses $n\mathrm{log}(n)·(\mathrm{log}(n)+1)/4$ comparator gates. \newline \newline
\textbf{Odd-Even Merge sort :} Odd-Even merge sort requires  $n(\mathrm{log}^{2}(n)-1)/2 + 1 $ comparator gates.\newline
We note that the sorting programs used do not increase the size of the circuit operating on encrypted input bit sequences.

\section{Experimental evaluation \\ and results}
\label{Sec:Eval}
In this we execute the implementations of the programs above, and we measure their wall time and compare this cost with the theoretical complexity depending on the sizes of the input data.
The figure \autoref{fig:image_sf7} represents minumum and maximum wall time of \texttt{XOR} done on random inputs. We can see that the wall time increase lineary with number of bits which confirm the theoretical cost.
\begin{figure}[!H]%on ouvre l'environnement figure
\centering
\includegraphics[width=7cm, height=6cm]{f4.pdf} 
\caption{Wall time for XOR} 
\label{fig:image_sf0} %l'étiquette pour faire référence à cette image
\end{figure}

We notice the same complexity for both \texttt{Bit Majority} and \texttt{sum of integers}\autoref{fig:image_sf1}.
\begin{figure}[!H]%on ouvre l'environnement figure
\centering
\includegraphics[width=6cm, height=5cm]{f8.pdf} 
\caption{Wall time for sum} 
\label{fig:image_sf1} %l'étiquette pour faire référence à cette image
\end{figure}

And we confirm as well the $O(n^{3})$ complexity for the \texttt{matrix product}\autoref{fig:image_sf2}.
\begin{figure}[!H]%on ouvre l'environnement figure
\centering
\includegraphics[width=9cm, height=6cm]{f9.pdf} 
\caption{Wall time for matrix product} 
\label{fig:image_sf2} %l'étiquette pour faire référence à cette image
\end{figure}
